---
title: "Getting Started with RMediation"
author: "RMediation Development Team"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with RMediation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
)
library(RMediation)
```

## Introduction

The RMediation package provides rigorous statistical methods for mediation analysis in observational and experimental designs. It addresses the known limitations of normal-theory confidence intervals (e.g., Sobel test) by implementing advanced methods that account for the non-normal distribution of the indirect effect.

## Basic Usage

### Single Mediator Analysis

To compute a confidence interval for a single mediator using summary statistics:

```{r example1}
# Example: Single mediator with known coefficients and standard errors
result <- medci(
    mu.x = 0.5, # Effect of X on M
    mu.y = 0.6, # Effect of M on Y (controlling for X)
    se.x = 0.08, # Standard error of a path
    se.y = 0.04, # Standard error of b path
    rho = 0, # Correlation between a and b (often 0 for experimental X)
    type = "dop" # Distribution of the product method
)

# Display the structure of the result
str(result)

# Or access specific components
cat("Confidence Interval:", result$`95% CI`, "\n")
cat("Estimate:", result$Estimate, "\n")
cat("Standard Error:", result$SE, "\n")
```

### Multiple Mediators

For more complex mediation models with multiple mediators:

```{r example2}
# Example: Two sequential mediators
result2 <- ci(
    mu = c(b1 = 1, b2 = .7, b3 = .6, b4 = .45),
    Sigma = c(.05, 0, 0, 0, .05, 0, 0, .03, 0, .03),
    quant = ~ b1 * b2 * b3 * b4,
    type = "MC",
    plot = FALSE # Set to TRUE to see visualization
)

str(result2)
```

### Using S7 Classes

The package includes modern S7 object-oriented classes:

```{r example3, eval = FALSE}
# Create a ProductNormal distribution object
pn <- ProductNormal(
    mu = c(0.5, 0.3), # Means of the two normal variables
    Sigma = matrix(c(0.01, 0.002, 0.002, 0.01), 2, 2) # Covariance matrix
)

# Compute confidence interval
ci_result <- ci(pn, level = 0.95)

show(pn) # Use show instead of print
str(ci_result) # Use str to show structure
```

## Available Methods

RMediation offers several methods for computing confidence intervals:

- `"dop"`: Distribution of the product (default for medci)
- `"MC"`: Monte Carlo method
- `"asymp"`: Asymptotic normal approach
- `"all"`: All methods

```{r methods}
# Compare different methods
comparison <- medci(
    mu.x = 0.5,
    mu.y = 0.6,
    se.x = 0.08,
    se.y = 0.04,
    rho = 0,
    type = "all",
    plot = FALSE
)

# Show comparison
str(comparison)
```

## Integration with Structural Equation Modeling

RMediation integrates seamlessly with popular SEM packages:

```{r lavaan-integration, eval = FALSE}
library(lavaan)

# Define a simple mediation model
model <- '
  # Direct effect
  Y ~ c*X

  # Mediator
  M ~ a*X
  Y ~ b*M

  # Indirect effect
  ab := a*b
  # Total effect
  total := c + (a*b)
'
# Simulate data
set.seed(123)
n <- 1000
X <- rnorm(n)
M <- 0.5 * X + rnorm(n)
Y <- 0.3 * M + 0.2 * X + rnorm(n)
df <- data.frame(X, M, Y)

# This would be fitted with your data
fit <- sem(model, data = df)

# Automatically compute CI for defined parameters
ci(fit) # This would auto-detect 'ab' parameter
```

## Validation and Error Handling

The package includes robust validation for inputs:

```{r validation, eval = FALSE}
# These would throw helpful error messages:
# medci(mu.x = 0.5, mu.y = 0.6, se.x = -0.1, se.y = 0.04)  # Invalid negative SE
# ci(mu = c(0.5), Sigma = matrix(1), quant = ~ b1)  # Dimension mismatch
```

## Conclusion

RMediation provides a comprehensive solution for mediation analysis with proper statistical inference. The package offers multiple methods for computing confidence intervals, seamless SEM integration, and robust error handling.

For more advanced usage and method comparisons, see our other vignettes.
